### Fitting  the model with full features analysed;
* DPP4 is the treatment, and hba1c is still kept as the outcome
* ncurrtx and drugline are treated as categorical variables.
* Orthongonalization implelemented
* Missingness allowed.

```{r}
# removing all the variables from the environment when re-running the analysis.
rm(list = ls(all.names = T))
# Setting seed to ensure same result are reproduced each time.
set.seed(1)
```
### loading the dataset into the workingt global environment.
```{r}
load("/slade/CPRD_data/mastermind_2019/Samuel/cprd_19_HDSproject_cohort.Rdata")
```
### loading the required library for the data analysis and manipulations.
```{r echo=FALSE, results='hide'}
library (grf, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/grf')
library (tidyverse, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/')
library (mice, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/')
library (sufrep, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/')
library (caret, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/')
library(DiagrammeR, lib.loc = '/slade/home/ns650/R/x86_64-pc-linux-gnu-library/4.0/')
```
### Peeking thge head of the data
```{r}
head(data)
```
```{r}
feature.filter <- c("datedrug", "drugstopdate", "MFN", "SU", 
                    "Acarbose", "GLP1", "Glinide", "DPP4", "INS", "SGLT2", "TZD",
                    "bestdiagdate", "ethcode", "posthba1c6mmmol", "posthba1c12mmmol",
                    "eth5", "eth16", "yrdrugstart", "ethnicitylong", "drugcombo")
```
```{r}
data %>% select(fldrug) %>% unique()
```
```{r}
stop_drug <- data %>% select(contains("stopdrug")) %>%
  names()
stop_drug
```

```{r}
all_col <- c(names(data))
longfeatlist <- all_col[which(!all_col %in% feature.filter)]
longfeatlist <- longfeatlist[which(!longfeatlist %in% stop_drug)]
longfeatlist <- c(longfeatlist, "stopdrug12m_3mFU")
longfeatlist
```
```{r}
sapply(data[, longfeatlist], class)
```
#### picking the char column band factor columns from the above feature list.
```{r}
required_cat_col <- c(
  "malesex",
  "Category", # character
  "fldrug",
  "drugcombo",
  "drugclass",
  "drugline",
  "ncurrtx"
)
```

```{r}
numeric_col <- longfeatlist[!longfeatlist %in% required_cat_col]
numeric_col
```

#### Extracting the drug cohorts separately and combining them with a balanced number of rows to check if it will impact the average treatment effetcs values
```{r}
# # Extracting cohorts on DPP4
DPP4.cohorts <- data %>%
  select("pateddrug":"fladherence_t") %>%
  group_by(pateddrug, drugclass) %>%
  filter(drugclass == "DPP4")
 # Extracting cohort on SGLT2
SGLT2.cohorts <- data %>%
  select("pateddrug":"fladherence_t") %>%
  group_by(pateddrug, drugclass) %>%
  filter(drugclass == "SGLT2")

# Balancing the sample, by selecting equal rows from DPP4 to match the sample amount in SGLT2
sglt2.dpp4 <- rbind(SGLT2.cohorts, DPP4.cohorts)

```
```{r}
nrow(sglt2.dpp4) 
# 16425
nrow(SGLT2.cohorts)
# 16424
nrow(DPP4.cohorts)
# 50608
#nrow(DPP4.cohorts[0:nrow(SGLT2.cohorts), ])
```

```{r}
sglt2.dpp4 <- sglt2.dpp4[, which(names(sglt2.dpp4) %in% longfeatlist)]
sglt2.dpp4 <- sglt2.dpp4 %>% rename("stopdrug" = "stopdrug12m_3mFU")
sglt2.dpp4 <- within(sglt2.dpp4, drugclass <- factor(drugclass))
sglt2.dpp4$Category <- as.factor(sglt2.dpp4$Category)
sglt2.dpp4$malesex <- as.integer(sglt2.dpp4$malesex)
sglt2.dpp4 <- within(sglt2.dpp4, deltaHba1c <- posthba1c_final - prehba1cmmol)
sglt2.dpp4 <- sglt2.dpp4[which(complete.cases(sglt2.dpp4$stopdrug)), ]
sglt2.dpp4$W <- if_else(sglt2.dpp4$drugclass == "DPP4", 1, 0)
sglt2.dpp4 <- sglt2.dpp4[sample(1:nrow(sglt2.dpp4)), ]

sglt2.dpp4 %>%
  head()
```
### Splitting the samples into train and test sample for evaluation later
```{r}
dftd_train <- sglt2.dpp4 %>% group_by(drugclass) %>%
  sample_frac(.6)
dftd_train <- data.frame(dftd_train)
#dftd_test <- sglt2.dpp4[!(sglt2.dpp4$pateddrug %in% dftd_train$pateddrug), ]
dftd_test <- subset(sglt2.dpp4, !(pateddrug %in% dftd_train$pateddrug))
dftd_train$pateddrug <- NULL ; dftd_test$pateddrug <- NULL
prop.table(table(dftd_train$drugclass))
prop.table(table(dftd_test$drugclass))
```

### Preprocessing for the training set.
```{r}
# setting a global option for  na
options(na.action = "na.pass")
# setting Y, and W
Y <- dftd_train$stopdrug
W <- dftd_train$W



X_train <- data.frame(dftd_train) %>% 
  select(-W, -stopdrug, -deltaHba1c, -drugclass, -posthba1c_final) 

dlx <- model.matrix(~ 0 + ., data = X_train["drugline"])
ncx <- model.matrix(~ 0 + ., data = X_train["ncurrtx"])
cax <- model.matrix(~ 0 + ., data = X_train["Category"])
fldx <- model.matrix(~ 0 + ., data = X_train["fldrug"])

# binding the data set
X_train <- cbind(X_train[, -which(names(X_train) %in% c("drugline", "ncurrtx", "Category", "fldrug"))], dlx, ncx, cax, fldx)
X_train <- X_train %>%model.matrix(~ ., data = .)

X_train <- X_train[, -1]

#head(X_train)
any(is.na(X_train))
```

```{r}
W.forest <- regression_forest(X = X_train, Y = W,
                              num.trees = 5000,
                              ci.group.size = 2,
                              min.node.size = 10,
                              alpha = 0.05,
                              imbalance.penalty = 0.05,
                              honesty = TRUE,
                              tune.parameters = "all")

W.hat <- predict(W.forest)$predictions

```
```{r}
df_Y <- data.frame(dftd_train) %>%
  select(-W, -stopdrug, -posthba1c_final, -drugclass, -ncurrtx, -drugline, -fldrug, -flt2dmduration) %>%
  model.matrix(~ ., data = .)

any(is.na(df_Y))
# Removing the intercept
df_Y <- df_Y[, -1]
head(df_Y)
Y.forest <- regression_forest(X = df_Y, Y = Y,
                              num.trees = 5000,
                              ci.group.size = 2,
                              min.node.size = 10,
                              alpha = 0.05,
                              imbalance.penalty = 0.05,
                              honesty = TRUE,
                              tune.parameters = "all")

Y.hat <- predict(Y.forest)$predictions
```
```{r}
hist(W.hat, col = "pink")

```
```{r}
hist(Y.hat, col = "green")
```
fitting the model first setup
```{r}
cf1 <- causal_forest(X = df_Y, Y = Y,
                    W = W,
                    Y.hat = Y.hat,
                    W.hat = W.hat,
                    num.trees = 5000,
                    min.node.size = 10,
                    ci.group.size = 2,
                    honesty = TRUE,
                    tune.parameters = "all")

```
fitting the model second setup
```{r}
cf2 <- causal_forest(X = df_Y, Y = Y,
                    W = W,
                    Y.hat = Y.hat,
                    W.hat = W.hat,
                    num.trees = 7000,
                    min.node.size = 10,
                    ci.group.size = 2,
                    honesty = TRUE,
                    tune.parameters = "all")

```
fitting the model third setup
```{r}
cf3 <- causal_forest(X = df_Y, Y = Y,
                    W = W,
                    Y.hat = Y.hat,
                    W.hat = W.hat,
                    num.trees = 5000,
                    min.node.size = 10,
                    ci.group.size = 2,
                    sample.fraction = 0.5,
                    honesty = TRUE,
                    honesty.fraction = 0.5,
                    honesty.prune.leaves = TRUE,
                    mtry = ncol(df_Y),
                    tune.parameters = c("min.node.size", "honesty.prune.leaves"))

```
fitting the model fourth setup
```{r}
cf4 <- causal_forest(X = df_Y, Y = Y,
                    W = W,
                    Y.hat = Y.hat,
                    W.hat = W.hat,
                    num.trees = 7000,
                    min.node.size = 10,
                    ci.group.size = 2,
                    sample.fraction = 0.5,
                    honesty = TRUE,
                    honesty.fraction = 0.5,
                    mtry = ncol(df_Y),
                    tune.parameters = c("min.node.size", "honesty.prune.leaves"))

```

### Preprocessing for the test set
```{r}
# factor = drug class, drug line, ncurrtx

Y.test <- dftd_test$stopdrug
W.test <- dftd_test$W

X_test <- data.frame(dftd_test) %>% 
  select(-W, -stopdrug, -deltaHba1c, -posthba1c_final, -drugclass)

dltx <- model.matrix(~ 0 + ., data = X_test["drugline"])
nctx <- model.matrix(~ 0 + ., data = X_test["ncurrtx"])
#dctx <- model.matrix(~ 0 + ., data = X_test["drugcombo"])
catx <- model.matrix(~ 0 + ., data = X_test["Category"])
fldtx <- model.matrix(~ 0 + ., data = X_test["fldrug"])

X_test <- cbind(X_test[, -which(names(X_test) %in% c("drugline", "ncurrtx", "drugcombo", "Category", "fldrug"))], dltx, nctx, catx, fldtx)

X_test <- X_test %>% model.matrix(~ ., data = .)

any(is.na(X_test))
X_test <- X_test[, -1]
#head(X_test)

```

### Fitting the Evaluation Forest.
```{r}
W.test.hat <- regression_forest(X = X_test, Y = W.test,
                                num.trees = 2000,
                                ci.group.size = 2,
                                min.node.size = 10,
                                honesty = TRUE,
                                tune.parameters = "all")
W.T.hat <- predict(W.test.hat)$predictions
## Replicating for Y. test
df_Yt <- data.frame(dftd_test) %>%
  select(-W, -stopdrug, -posthba1c_final, -drugline, -ncurrtx, -drugclass, -fldrug, -flt2dmduration) %>%
  model.matrix(~ ., data = .)
df_Yt <- df_Yt[, -1]
## Fitting the y.test model
Y.test.hat <- regression_forest(X = df_Yt, Y = Y.test,
                                num.trees = 2000,
                                ci.group.size = 2,
                                min.node.size = 10,
                                honesty = TRUE,
                                tune.parameters = "all")

Y.T.hat <- predict(Y.test.hat)$predictions

eval.forest <- causal_forest(X = df_Yt, Y = Y.test,
                             W = W.test,
                             W.hat = W.T.hat,
                             Y.hat = Y.T.hat,
                             num.trees = 2000,
                             alpha = 0.05,
                             imbalance.penalty = 0.05,
                             ci.group.size = 2,
                             honesty = TRUE,
                             tune.parameters = "all")
```
## Checking the forest calibrations for both the trained forest and the test forest
```{r}
test_calibration(eval.forest, vcov.type = "HC0")
```
```{r}
test_calibration(cf1,vcov.type = "HC0")
```
```{r}
test_calibration(cf2,vcov.type = "HC0")
```
```{r}
test_calibration(cf3,vcov.type = "HC0")
```
```{r}
test_calibration(cf4,vcov.type = "HC0")
```
```{r}
average_treatment_effect(cf4, target.sample = "overlap")
```

```{r}
tau.hat <- predict(cf4, df_Yt)$predictions
```
```{r}
TOC <- rank_average_treatment_effect(eval.forest, tau.hat, target = "AUTOC")
plot(TOC, col = "red");
```
### Getting a point estimate using the trained forest
```{r}
cf_pred <- predict(cf4, df_Yt, estimate.variance = TRUE)
```
```{r}
hist(cf_pred$predictions, main="Causal forests: Point estimate of CATE",
     xlab = "predict oob cate", ylab = "number of patients")
abline(v=mean(cf_pred$predictions), col = "green", lty = "dashed", lwd=3)
abline(v=0, col = "red", lty = "dashed", lwd = 3)
```
### Estimating the average treatment effect
```{r}
ate.hat <- average_treatment_effect(cf4, target.sample = "control")
print(paste("95% CI for ATE:", round(ate.hat["estimate"], 3), "+/-", round(1.96*ate.hat["std.err"], 3)))
```
```{r}
ate.hat <- average_treatment_effect(cf4, target.sample = "treated")
print(paste("95% CI for ATE:", round(ate.hat["estimate"], 3), "+/-", round(1.96*ate.hat["std.err"], 3)))
```
```{r}
ate.hat <- average_treatment_effect(cf4, target.sample = "overlap")
print(paste("95% CI for ATE:", round(ate.hat["estimate"], 3), "+/-", round(1.96*ate.hat["std.err"], 3)))
```

```{r}
oob_predictions <- predict(cf4, estimate.variance = TRUE)
Z.scores <- oob_predictions$predictions / oob_predictions$variance.estimates
```
### Visualizing the Z scores
The Z.score measures significance of the heterogeneous treatment effect estimates. Under the null, we might expect the Z-scores to be roughly standard Gaussian.
```{r}
hist(Z.scores)
```
### Anothervisualization
```{r}
plot(oob_predictions$predictions, sqrt(oob_predictions$variance.estimates))
abline(0, 1/1.96); abline(0, -1/1.96)
```
It looks like the Z-scores have a noted left skew. It clear that we have some individuals for whom tau is negative, and also have evidence that there are some people for whom tau is positive.
### plotting first prediction on the second predictions
```{r}
# plot(oob2_predictions$predictions, oob_predictions$predictions); abline(0,1)
# summary(lm(oob_predictions$predictions ~ oob2_predictions$predictions))
```

```{r}
oob_cf <- oob_predictions$predictions
varimp <- variable_importance(cf4, decay.exponent = 2, max.depth = 4)
selected.idx <- which(varimp > 0.05)
```
```{r}
names(varimp) <- colnames(cf4$X.orig)
sorted_varimp <- sort(varimp, decreasing = T)
varim_abovemean <- sorted_varimp[selected.idx]
t(varim_abovemean)
```
```{r}
barplot(varim_abovemean, las = 2, col = "red", main = "Variable importance plot")
```





